---
title: "Introduction to finalwork"
author: "Xinyu Li"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to finalwork}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

__finalwork__ is an R package that implements part of the simulation section of a nonlinear missing data completion algorithm called GFM-TW. It includes two Poisson data simulation generation functions ( __DGP1__ , __DGP2__ ), which are used to generate Poisson data without and with intercepts, a function ( __DGP_miss__ ) that simulates the missing data algorithm, and a Poisson version of the GFM-TW algorithm ( __GFM_TW__ ), which is used to solve the problem of Poisson data completion with missing values, and a Rcpp function ( __RMSEC__ ) to calculate the RMSE indicator.

## DGP functions

The function __DGP1__ is used to simulate the random number generation process of the following data:   
$$
\begin{aligned}
X_{ij}&\sim Pois(exp(z_{ij})),\ z_{ij}=\alpha_{i}^{\prime}\gamma_{j},i=1,\dots,N,\ j=1,\dots,P  \\
\alpha_{i}&=(\alpha_{i1},1,\alpha_{i2})^{\prime},\ \alpha_{i1}\sim U(0,1),\ \alpha_{i2}\sim N_{R_2}(0,I_{R_2})\\
\gamma_{j}&=(1,\gamma_{j1},\gamma_{j2})^{\prime},\ \gamma_{j1}\sim U(0,1),\ \gamma_{j2}\sim N_{R_2}(0,I_{R_2}) \\
\end{aligned}
$$

The source R code for _DGP1_ is as follows:   

```{r,eval=FALSE}
DGP1<-function(N,P,R2){
  X<-matrix(rep(0,N*P),nrow=N)
  #alpha~(U(0,1),1,alphai2)
  Alpha<-cbind(runif(N),rep(1,N),mvrnorm(N,rep(0,R2),Sigma=diag(R2))) 
  #gamma~(1,U(0,1),alphai2)
  Gamma<-cbind(rep(1,P),runif(P),mvrnorm(P,rep(0,R2),Sigma=diag(R2))) 
  Z<-exp(Alpha%*%t(Gamma))
  X<-matrix(rpois(N*P,lambda=Z),N,P)
  return(X)
}
```

The above code simulates generating Poisson data without intercept terms with $R_2+2$ factors.  

The function __DGP2__ is used to simulate the random number generation process of the following data:   

$$
\begin{aligned}
X_{ij}&\sim Pois(exp(z_{ij})),\ z_{ij}=\alpha_{i}^{\prime}\gamma_{j}+\mu_{j}  \\
\alpha_{i}&=(\alpha_{i1},\dots,\alpha_{iR_2}),\alpha_{ik}\sim N(0,0.5^2), \ i.i.d.,k=1,\dots,R_2\\
\gamma_{j}&\sim N_{R_2}(1,I_{R_2}) \\
\mu_{j}&\sim 0.5U(0,1),j=1,\dots,p
\end{aligned}
$$

The source R code for _DGP2_ is as follows:   

```{r,eval=FALSE}
DGP2<-function(N,P,R2){
  X<-matrix(rep(0,N*P),nrow=N)
  Alpha<-matrix(rnorm(N*R2,0,0.5),nrow=N)
  Gamma<-mvrnorm(P,rep(1,R2),Sigma=diag(R2))
  mu<-matrix(0.5*runif(P),N,P,byrow = TRUE)
  Z<-exp(Alpha%*%t(Gamma)+mu)
  X<-matrix(rpois(N*P,lambda=Z),N,P)
  return(X)
}
```

The above code simulates generating Poisson data with intercept terms with $R_2$ factors.   

## Missing mechanism setting

I set the missing mechanism to be completely random, but the number of missing rows and columns is fixed, denoted by $n_m$ and $p_m$. Specifically, I scrambled the row and column indexes of the data and extracted the $n_m$ and $p_m$ columns as missing rows and columns, and set random missing in the $n_m \times p_m$ submatrix composed of these rows and columns. The total missing data is about $1/4 n_m \times p_m$.

The source R code for _DGP_miss_ is as follows:   

```{r,eval=FALSE}
DGP_miss<-function(Nm,Pm,X){
  N<-nrow(X);P<-ncol(X)
  row_index<-sample(1:N,Nm,replace=FALSE)
  col_index<-sample(1:P,Pm,replace=FALSE)
  X_miss<-X 
  for(rindex in row_index){ # Construct column missing  
    num<-sample(1:(Pm/4),size=1,replace=FALSE) #the num of missing columns in row rindex
    col_missindex<-sample(col_index,size=num,replace = FALSE)
    X_miss[rindex,col_missindex]<-NA
  }
  for(cindex in col_index){ # Construct row missing
    num<-sample(1:(Nm/4),size=1,replace=FALSE) #the num of missing rows in column cindex
    row_missindex<-sample(row_index,size=num,replace = FALSE)
    X_miss[row_missindex,cindex]<-NA
  }
  
  return(X_miss)   
}
```

## GFM-TW algorithm  

The following function is used to implement our proposed nonlinear missing data completion algorithm. Here I only show the part of the algorithm for Poisson data.

```{r,eval=FALSE}
GFM_TW<-function(X_miss,num_factor,type="poission",
                 detail="no intercept"){
  N<-nrow(X_miss);P<-ncol(X_miss)
  #mark matrix,mij=1(TRUE) denotes missing
  missing <- is.na(X_miss)        
  #row index with non missing values
  goodN <- rowSums(is.na(X_miss)) == 0   
  #column index with non missing values
  goodP <- colSums(is.na(X_miss)) == 0
  X_TALL<-X_miss[,goodP]
  X_WIDE<-X_miss[goodN,]
  
  #data is the type of poisson
  if(type=="poisson"){
    if(detail=="no intercept"){
      #use TALL to estimate factors
      gfm_TALL <- gfm(list(X_TALL), 'poisson',  q=num_factor, verbose = FALSE)
      H_hat1<-gfm_TALL$hH
      #use WIDE to estimate loadings and intercept
      gfm_WIDE <- gfm(list(X_WIDE), 'poisson',  q=num_factor, verbose = FALSE)
      B_hat1<-gfm_WIDE$hB
      mu_hat1<-gfm_WIDE$hmu
      Mu_hat1<-t(matrix(rep(mu_hat1,N),ncol=N))
      
      #the first estimate, take the rounded value of the mean as an estimate
      X_hat1<-matrix(rep(0,N*P),nrow=N)
      X_hat1<-H_hat1%*%t(B_hat1)+Mu_hat1
      X_hat1<-round(exp(X_hat1))
      #impute the matrix
      X_hat_tmp<-matrix(rep(t(X_miss),1),ncol=P,byrow=TRUE)
      X_hat_tmp[missing]<-X_hat1[missing]
      
      #reestimate
      gfm_ALL<-gfm(list(X_hat_tmp), 'poisson',  q=num_factor, verbose = FALSE)
      H_hat2<-gfm_ALL$hH
      B_hat2<-gfm_ALL$hB
      mu_hat2<-gfm_ALL$hmu
      Mu_hat2<-t(matrix(rep(mu_hat2,N),ncol=N))
      
      X_hat2<-matrix(rep(0,N*P),nrow=N)
      X_hat2<-H_hat2%*%t(B_hat2)+Mu_hat2
      X_hat2<-round(exp(X_hat2))
      X_complete<-matrix(rep(t(X_miss),1),ncol=P,byrow=TRUE)
      X_complete[missing]<-X_hat2[missing]
    }
    if(detail=="intercept"){
      #use TALL to estimate factors
      gfm_TALL <- gfm(list(X_TALL), 'poisson',  q=num_factor, verbose = FALSE)
      H_hat1<-gfm_TALL$hH
      #use WIDE to estimate loadings and intercept
      gfm_WIDE <- gfm(list(X_WIDE), 'poisson',  q=num_factor, verbose = FALSE)
      B_hat1<-gfm_WIDE$hB
      mu_hat1<-gfm_WIDE$hmu
      Mu_hat1<-t(matrix(rep(mu_hat1,N),ncol=N))
      
      #the first estimate, take the rounded value of the mean as an estimate
      X_hat1<-matrix(rep(0,N*P),nrow=N)
      X_hat1<-H_hat1%*%t(B_hat1)+Mu_hat1
      X_hat1<-round(exp(X_hat1))
      X_hat_tmp<-matrix(rep(t(X_miss),1),ncol=P,byrow=TRUE)
      X_hat_tmp[missing]<-X_hat1[missing]
      
      #reestimate
      gfm_ALL<-gfm(list(X_hat_tmp), 'poisson',  q=num_factor, verbose = FALSE)
      H_hat2<-gfm_ALL$hH
      B_hat2<-gfm_ALL$hB
      mu_hat2<-gfm_ALL$hmu
      Mu_hat2<-t(matrix(rep(mu_hat2,N),ncol=N))
      
      X_hat2<-matrix(rep(0,N*P),nrow=N)
      X_hat2<-H_hat2%*%t(B_hat2)+Mu_hat2
      X_hat2<-round(exp(X_hat2))
      X_complete<-matrix(rep(t(X_miss),1),ncol=P,byrow=TRUE)
      X_complete[missing]<-X_hat2[missing]
    }
  }
  
  #return the complete matrix
  return(X_complete)
}
```

## Evaluation indicator

The following is a function for calculating the RMSE metric using Rcpp:

```{r,eval=FALSE}
double RMSEC(const NumericMatrix& Xhat, 
             const NumericMatrix& X,int missnum) {
  double sum = 0;
  int nrow = X.nrow();
  int ncol = X.ncol();
  double rmse=0;
  double diff=0;
  for (int i = 0; i < nrow; i++) {
    for (int j = 0; j < ncol; j++) {
      diff = X(i, j) - Xhat(i, j);
      sum += diff * diff;
    }
  }
  rmse=std::sqrt(sum / missnum);
  return rmse;
}
```

The definition of RMSE is    
$$
\operatorname{RMSE}=\sqrt{\frac{1}{|\mathcal{D}|-\left|\mathcal{D}_{o}\right|} \sum_{(i, j) \in \mathcal{D} \mathcal{D}_{o}}\left(\widehat{X}_{i j}-X_{i j}\right)^{2}},
$$

where $\widehat{X}_{i j}$ is the estimate in the completed matrix $X_{i j}$, $\mathcal{D}_{o}$ is a set containing index pairs of all observed variables, and $|\cdot|$ represents the number of elements in the set.


## Performance comparison  

The following newly defined function compares the effect of GFM-TW and TW algorithms (TW-sd and TW-raw, the former use the standardized data and the later use raw data.) on the basis of generating the same data matrix, and returns the imputed matrix obtained by different algorithms and the number of missing values.

```{r,eval=TRUE}
library(finalwork)
library(fbi)
compare_method<-function(N,P,Nm,Pm,itermax,R2,types='poisson',details="no intercept"){
  #generate a list to reserve the result of each iter
  #list[[iter]] contains X_true,missnum,X_complete for three methods
  result<-list()
  for(iter in 1:itermax){
    if(types=="poisson"){
      if(details=="no intercept"){
        #generate data matrix and artificially
        #creating missing values
        X_true<-DGP1(N,P,R2)
        X_true_miss<-DGP_miss(Nm,Pm,X_true)
        missnum<-sum(is.na(X_true_miss))
        ##GFM-TW
        X_complete_GFMTW<-GFM_TW(X_true_miss,R2+2,types,details)
        ##TW-sd
        X_complete_TWsd<-tw_apc(X_true_miss,R2+2,center=TRUE,standardize = TRUE,re_estimate = TRUE)$data
        X_complete_TWsd<-round(X_complete_TWsd)
        ##TW-raw
        X_complete_TWraw<-tw_apc(X_true_miss,R2+2,center=FALSE,standardize = FALSE,re_estimate = TRUE)$data
        X_complete_TWraw<-round(X_complete_TWraw)
      }
      if(details=="intercept"){
        X_true<-DGP2(N,P,R2)
        X_true_miss<-DGP_miss(Nm,Pm,X_true)
        missnum<-sum(is.na(X_true_miss))
        ##GFM-TW
        X_complete_GFMTW<-GFM_TW(X_true_miss,R2,types,details)
        ##TW-sd
        X_complete_TWsd<-tw_apc(X_true_miss,R2,center=TRUE,standardize = TRUE,re_estimate = TRUE)$data
        X_complete_TWsd<-round(X_complete_TWsd)
        ##TW-raw
        X_complete_TWraw<-tw_apc(X_true_miss,R2,center=FALSE,standardize = FALSE,re_estimate = TRUE)$data
        X_complete_TWraw<-round(X_complete_TWraw)
      }
    }
    #reserve results
    result[[iter]]<-list(X_true=X_true,missnum=missnum,
                         X_complete_GFMTW=X_complete_GFMTW,
                         X_complete_TWsd=X_complete_TWsd,
                       X_complete_TWraw=X_complete_TWraw)
  }
  return(result)
}
```

The following newly defined function reports the RMSE result.    

```{r,eval=TRUE}
library(finalwork)
report_RMSE<-function(res,iternum){
  RMSE_result<-matrix(0,nrow=iternum,ncol=3)
  for(iter in 1:iternum){
  RMSE_result[iter,1]<-RMSEC(res[[iter]]$X_complete_GFMTW,res[[iter]]$X_true,res[[iter]]$missnum)
  RMSE_result[iter,2]<-RMSEC(res[[iter]]$X_complete_TWsd,res[[iter]]$X_true,res[[iter]]$missnum)
  RMSE_result[iter,3]<-RMSEC(res[[iter]]$X_complete_TWraw,res[[iter]]$X_true,res[[iter]]$missnum)
  }
  return(RMSE_result)
}
```

## Examples

### Example1   

In this example, we simulate the whole process of missing completion of Poisson data without intercepts. The reported matrix is a $M \times 3$ matrix, where $M$ is the iteration number.

```{r,eval=TRUE}
#setting
library(finalwork)
library(fbi)
iter1=10
N1=200;P1=300;R1=1;Nm1<-80;Pm1<-80
#simulation
result1<-compare_method(N1,P1,Nm1,Pm1,iter1,R1,'poisson','no intercept')

#report
report_result<-report_RMSE(result1,iter1)
report_result
```

### Example2   

In this example, we simulate the whole process of missing completion of Poisson data with intercepts. The reported matrix is a $M \times 3$ matrix, where $M$ is the iteration number.

```{r,eval=TRUE}
#setting
iter2=10
N2=200;P2=300;R=3;Nm2<-80;Pm2<-80
#simulation
result1<-compare_method(N2,P2,Nm2,Pm2,iter2,R,'poisson','intercept')
#report
report_result<-report_RMSE(result1,iter2)
report_result
```